Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-20 Draft.Rmd
edited format of boosting code

origin %>% ggplot(aes(x=lat, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration (mu g/m^3)", x="Latitude (degrees)", y="PM2.5")

origin %>% ggplot(aes(x=value, fill=state)) + geom_histogram() + facet_wrap(~state) + 
  labs(title="PM2.5 Concentration (mu g/m^3)", x="State", y="PM2.5") 
```

@@

```{r Bootstrap test}
model3 = boost_tree( mtry = integer(), trees = integer(), min_n = integer(),
  tree_depth = integer(), learn_rate = numeric(), loss_reduction = numeric(),
  sample_size = numeric(), stop_iter = integer()) %>%
  set_engine("xgboost") %>%
  set_mode("regression") %>%
  translate()

# preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value~lon+lat, data = training(dat_split)) %>%
  # convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

folds3 <- 
  recipes::bake(
    preprocessing_recipe, 
    new_data = training(dat_split)
  ) %>%  
  rsample::vfold_cv(v = 5)

@@

# grid specification
xgboost_params <- 
  dials::parameters(min_n(), tree_depth(), learn_rate(), loss_reduction())

xgboost_grid <- dials::grid_max_entropy(xgboost_params, size = 60)
knitr::kable(head(xgboost_grid))

@@

# hyperparameter tuning
xgboost_tuned <- tune::tune_grid(
  object = xgboost_wf,
  resamples = folds3,
  grid = xgboost_grid,
  metrics = yardstick::metric_set(rmse, rsq, mae),
  control = tune::control_grid(verbose = TRUE)
)

xgboost_tuned %>%
  tune::show_best(metric = "rmse") %>%
  knitr::kable()

@@

# Create Random Forest Model
set.seed(123)
# Use grid of tuning parameters
model4 <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")
# Create workflow
  wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model4)

# Fit the model on the full training dataset using the `fit()` function
res4 <- fit(wf4, data = train_scaled)
# Check performance on the complete training data
res4 %>% 
    extract_fit_engine() %>% 
    summary()
# Check performance using cross-validation
folds <- vfold_cv(train, v = 10)
folds
# Determine performance of model by resampling
res3 <- fit_resamples(wf4, folds) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res3 %>% 
    collect_metrics()

## Try a grid of tuning parameters
model <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")

wf <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model)

## Fit model over grid of tuning parameters
res <- tune_grid(wf, resamples = folds)
res %>% 
    collect_metrics()

>>>>

@@

```{r Boost}
# Preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value~lon+lat, data = training(dat_split)) %>%
  # convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

folds3 <- recipes::bake(preprocessing_recipe, new_data = training(dat_split)) %>%  
  rsample::vfold_cv(v = 5)

@@

# Grid specification
xgboost_params <- dials::parameters(min_n(), tree_depth(), learn_rate(), loss_reduction())

xgboost_grid <- dials::grid_max_entropy(xgboost_params, size = 60)

knitr::kable(head(xgboost_grid))

@@

# Hyperparameter tuning
# Takes a long time to run
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq, mae),
  control = tune::control_grid(verbose = TRUE))

# Find hyper parameter values which performed best at minimizing RMSE
xgboost_tuned %>%
  tune::show_best(metric = "rmse") %>%
  knitr::kable()
# Isolate the best performing hyperparameter values.
xgboost_best_params <- xgboost_tuned %>%
  tune::select_best("rmse")
knitr::kable(xgboost_best_params)

# Final boost model
xgboost_model_final <- xgboost_model %>% 
  finalize_model(xgboost_best_params)

# Eval. model performance on test data
train_processed <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction <- xgboost_model_final %>%
# fit the model on all the training data
  fit(formula = value ~ lon+lat, data = train_processed) %>%
  # predict the sale prices for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(ames_split))

xgboost_score_train <- train_prediction %>%
  yardstick::metrics(sale_price, .pred) %>%
  mutate(.estimate = format(round(.estimate, 2), big.mark = ","))

knitr::kable(xgboost_score_train)
```

@@

```{r Random Forests}
# Create Random Forest Model
model4 <- rand_forest(mtry = 5) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")

# Create workflow
  wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model4)
